---
title: "Analysis"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load relevant packages for pannel regression modelling
```{r}
library(tidyverse) # Modern data science library 
library(plm)       # Panel data analysis library
library(car)       # Companion to applied regression 
library(gplots)    # Various programing tools for plotting data
library(tseries)   # For timeseries analysis
library(lmtest)    # For hetoroskedasticity analysis
```




#Iniatal Data Anlsysis ggpairs()
```{r}

ggpairs_plot <- ggpairs(select(SS_imputed_data_final_logged, -Name, -Year, -'.imp', -'.id')) +
  theme_minimal()    

ggsave("ggpairs_plot.png", ggpairs_plot, width = 12, height = 6, dpi = 300)

```


#Inital Data Analysis specific for pannel data

```{r}

#coplot
pdf("myplot.pdf", width=10, height=10)
coplot(CO2 ~ Year|Name, type = "b", data= SS_imputed_data_final_logged)


#scatterplot

pdf("myplot2.pdf", width=10, height=10)
scatterplot(CO2 ~ Year|Name, data= SS_imputed_data_final_logged)


#---

#scaterplot for 10 random countries

set.seed(1234) # for reproducibility

# select a random sample of 10 'Name' values
random_names <- sample(unique(SS_imputed_data_final_logged$Name), 10)

# subset the data to include only the randomly selected 'Name' values
subset_data <- SS_imputed_data_final_logged[SS_imputed_data_final_logged$Name %in% random_names, ]

# create a scatterplot for each of the selected 'Name' values
pdf("myplot2a.pdf", width=12, height=12)
par(mfrow=c(2,5)) # arrange the plots in a 2x5 grid
for (name in random_names) {
  plot(subset_data$Year[subset_data$Name == name], 
       subset_data$CO2[subset_data$Name == name], 
       xlab = "Year", ylab = "CO2",
       main = paste("Scatterplot for", name))
  lines(subset_data$Year[subset_data$Name == name], 
        subset_data$CO2[subset_data$Name == name])
}
dev.off()




#OR



set.seed(123) # for reproducibility

# select a random sample of 10 'Name' values
random_names <- sample(unique(SS_imputed_data_final_logged$Name), 10)

# subset the data to include only the randomly selected 'Name' values
subset_data <- SS_imputed_data_final_logged[SS_imputed_data_final_logged$Name %in% random_names, ]

# create a scatterplot for each of the selected 'Name' values
myplot <- ggplot(subset_data, aes(x = Year, y = CO2, color = Name)) +
  geom_point() +
  facet_wrap(~Name, ncol = 5) +
  labs(title = "Scatterplot for Randomly Selected Names",
       x = "Year",
       y = "CO2")

ggsave("myplot2b.pdf", plot = myplot, width = 10, height = 10)





#OR



# select a random sample of 10 'Name' values
set.seed(123) # for reproducibility
random_names <- sample(unique(SS_imputed_data_final_logged$Name), 10)

# subset the data to include only the randomly selected 'Name' values
subset_data <- SS_imputed_data_final_logged %>%
  filter(Name %in% random_names)

# create a scaterplot with all the randomly selected 'Name' values
ggplot(subset_data, aes(x = Year, y = CO2, color = Name)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle("Scatterplot for 10 Random Countries") +
  theme_minimal()



```



#Heterogeniety
```{r}

# Heterogeneity across countries
pdf("myplot3.pdf", width=10, height=10)
plotmeans(CO2 ~ Name, data = SS_imputed_data_final_logged)


#Heterogeneity across years
pdf("myplot4.pdf", width=10, height=10)
plotmeans(CO2 ~ Year, data = SS_imputed_data_final_logged)
```


#Further Data Analysis


#Basic OLS model

```{r}

#Basic OLS model
#The basic OLS regression model does not consider heterogeneity across countries or across years
ols <- lm(CO2 ~ ., data = select(SS_imputed_data_final_logged, -c(Name, .imp, .id)))
summary(ols)

plot(ols)

#Visually represented
yhat <- ols$fitted
ggplot(select(SS_imputed_data_final_logged, -c(Name, .imp, .id)), aes(x = EU_Exp, y = CO2))+
  geom_point() +
  geom_smooth(method=lm)


```

#Country-Specific Fixed Effects using Dummy Variables (LSDV Model)

```{r}


fixed.dum <-lm(CO2 ~ EU_Exp + factor(Name) - 1, data = select(SS_imputed_data_final_logged, -c(Name, .imp, .id,)))
summary(fixed.dum)

pdf("lsdv.pdf", width=10, height=10)
sampled_names <- sample(unique(SS_imputed_data_final_logged$Name), 10)
subset_data <- subset(SS_imputed_data_final_logged, Name %in% sampled_names)
subset_model <- lm(CO2 ~ EU_Exp + factor(Name) - 1, data = subset_data)
yhat <- predict(subset_model, newdata = subset_data)
scatterplot(yhat ~ EU_Exp | Name, data = subset_data, xlab ="GS Merch Export", ylab ="yhat", boxplots = FALSE, smooth = FALSE)
abline(lm(CO2 ~ EU_Exp, data = subset_data), lwd=3, col="red")



#OR 
library(RColorBrewer)

pdf("lsdv.pdf", width=10, height=10)

sampled_names <- sample(unique(SS_imputed_data_final_logged$Name), 10)

subset_data <- subset(SS_imputed_data_final_logged, Name %in% sampled_names)

subset_model <- lm(CO2 ~ EU_Exp + factor(Name) - 1, data = subset_data)

yhat <- predict(subset_model, newdata = subset_data)

palette <- brewer.pal(nlevels(subset_data$Name), "Set1") # create custom color palette

scatterplot(yhat ~ EU_Exp | Name, data = subset_data, xlab ="GS Merch Export", ylab ="yhat", boxplots = FALSE, smooth = FALSE, col=palette)

abline(lm(CO2 ~ EU_Exp, data = subset_data), lwd=3, col="red")

dev.off()



```






#Country-Specific Fixed Effects using the plm package
```{r}
fixed <- plm(CO2 ~ EU_Exp, data= select(SS_imputed_data_final_logged, -c(.imp, .id)), model="within")
summary(fixed)

fixef(fixed)

pFtest(fixed, ols)

#If the p-value is < 0.05 then the fixed effects model is a better choice, ours is almost equivalent to 0


```

# Random Effect modelling using plm

```{r}
random <- plm(CO2 ~ EU_Exp, data= select(SS_imputed_data_final_logged, -c(.imp, .id)), model="random")
summary(random)
```

#Fixed vs Random

```{r}

phtest(fixed, random)


# To decide between fixed or random effects you can run a Hausman test where the null hypothesis is that the preferred model is random effects vs. the alternative the fixed effects (see Green, 2008, chapter 9). It basically tests whether the unique errors are correlated with the regressors, the null hypothesis is they are not. If the p-value is significant (for example <0.05) then use fixed effects, if not use random effects. 

#Here it shows we should use the fixed effect model.
```




```{r}
#OLD BUT KEEP FOR NOW

# Perform linear regression analysis on each completed dataset
#all the R2 of all the model around 30-40 other than GDP which was 74.5
model <- plm(CO2_kt ~ GDP_cur_USD + factor(Name), data = completed_data_1, model = "within")
summary(model)

model1 <- plm(CO2_kt ~ Merchandise_Exported_to_EU_thousandsUSD + factor(Name), data = completed_data_1, model = "within")
model2 <- plm(CO2_kt ~ Merchandise_Exported_to_EU_thousandsUSD + factor(Name), data = completed_data_2, model = "within")
model3 <- plm(CO2_kt ~ Merchandise_Exported_to_EU_thousandsUSD + factor(Name), data = completed_data_3, model = "within")
model4 <- plm(CO2_kt ~ Merchandise_Exported_to_EU_thousandsUSD + factor(Name), data = completed_data_4, model = "within")
model5 <- plm(CO2_kt ~ Merchandise_Exported_to_EU_thousandsUSD + factor(Name), data = completed_data_5, model = "within")


# Combine estimates and standard errors using Rubin's rules
fit_pool <- pool(list(model1, model2, model3, model4, model5))

# Print pooled coefficients and standard errors
summary(fit_pool)

pool

# Pool model results using plm::pool()
pooled <- pool(melded)

# View pooled estimates and standard errors
summary(pooled)

qqnorm(model1)
qqline(model1)

plot(model1)

```

