---
title: "Appendix4b"
output: word_document
---




#Load Data
```{r}
completely_loaded_data <- read_excel("/Users/jamesduffy/Library/CloudStorage/OneDrive-TheUniversityofManchester/BASS/Modules/Year\ 3/Sem\ 2/Diss/Data/EU_Effect.xlsx")
```

## Explore missing data

# Analysis of patterns in missing data
```{r}

#Count of NAs
summary(completely_loaded_data)
options(max.print = 50000)


# Plot the missing data pattern

png("md_pattern.png", width = 2600, height = 2400)
md.pattern(completely_loaded_data, plot = TRUE, rotate.names = TRUE)
dev.off()


#Alternative ways of analyzing missing data

p <- vis_miss(completely_loaded_data)
ggsave("missingness_plot.pdf", plot = p)

gg_miss_upset(completely_loaded_data)

```



## Remove Democracy and Electricity from coal
```{r}
completely_loaded_data_2 <- select(completely_loaded_data, -'democ', -'Electricity_from_coal_%oftotal')

# Plot the missing data pattern

png("md_pattern_2.png", width = 2600, height = 2400)
md.pattern(completely_loaded_data_2, plot = TRUE, rotate.names = TRUE)
dev.off()
```

##Exploring how to deal with NAs

```{r}

#Remove the 48 NA's in our trade data
completely_loaded_data_3 <- completely_loaded_data_2[complete.cases(completely_loaded_data_2[, c(9, 14)]), ]

# Plot the missing data pattern

png("md_pattern_3.png", width = 2600, height = 2400)
md.pattern(completely_loaded_data_3, plot = TRUE, rotate.names = TRUE)
dev.off()

#tidy up col names
names(completely_loaded_data_3)[8] <- "Urban_Pop_perc_of_total"
names(completely_loaded_data_3)[11] <- "Electricity_from_renewables_perc_oftotal"
names(completely_loaded_data_3)[12] <- "Manufacturing_value_added_perc_ofGDP"
names(completely_loaded_data_3)[13] <- "FDI_netinflow_perc_ofGDP"


#Remove NAs if they appear all together

completely_loaded_data_4 <- completely_loaded_data_3[!(rowSums(is.na(completely_loaded_data_3[, c("Urban_Pop_perc_of_total", "GDP_cur_USD", "Electricity_from_renewables_perc_oftotal", "Manufacturing_value_added_perc_ofGDP", "FDI_netinflow_perc_ofGDP")])) == 5), ]

png("md_pattern_4.png", width = 2600, height = 2400)
md.pattern(completely_loaded_data_4, plot = TRUE, rotate.names = TRUE)
dev.off()

```


#Exploirng having all the NAs removed

```{r}

#Remove the all NAs
completely_loaded_data_0NA <- completely_loaded_data_3[complete.cases(completely_loaded_data_3[, c(8, 10, 11, 12, 13)]), ]


# Plot the missing data pattern

png("md_pattern_0NA.png", width = 2600, height = 2400)
md.pattern(completely_loaded_data_0NA, plot = TRUE, rotate.names = TRUE)
dev.off()

attach(completely_loaded_data_3)

#compare averages
summary(completely_loaded_data_3)
summary(completely_loaded_data_3)
summary(completely_loaded_data_0NA)
```

# Imputation of rest of NAs using MICE

```{r}
# Impute missing data using mice with the rf method, first remove irrelevant variables
completely_loaded_data_5 <- completely_loaded_data_4[, !(names(completely_loaded_data_4) %in% c("Country_Code", "IPCC_annex"))]

completely_loaded_data_5_mice = mice(completely_loaded_data_5, m=5, method= 'rf', maxit = 5)


# Extract each imputed dataset using complete()
completed_data_1 <- complete(completely_loaded_data_5_mice, 1)
completed_data_2 <- complete(completely_loaded_data_5_mice, 2)
completed_data_3 <- complete(completely_loaded_data_5_mice, 3)
completed_data_4 <- complete(completely_loaded_data_5_mice, 4)
completed_data_5 <- complete(completely_loaded_data_5_mice, 5)

#Compare the imputed values in each data set to the original data



summary(completely_loaded_data_4)
summary(completed_data_1)
summary(completed_data_2)
summary(completed_data_3)
summary(completed_data_4)
summary(completed_data_5)
summary(completley_loaded_data_0NA)


# Impute missing values in the original data set using the pooled model. This data set is 5 times larger than the original 

imputed_data <- complete(completely_loaded_data_5_mice, action = "long")


```



#Now have three datasets
#imputed_data= all 5 commpleted datasets
#completed_data1-5= iniduvdal completed datasets
#completely_loaded_data_5= Original dataset with NAs




#Logarithmic transformation of imputed data

```{r}

imputed_data_numeric <- select(imputed_data, -.imp:-Year)

# apply log() to all columns except 'a', 'b', and 'c'
imputed_data_numeric_logged <- imputed_data_numeric %>% 
  mutate_at(vars(-FDI_netinflow_perc_ofGDP, -Electricity_from_renewables_perc_oftotal, -Merchandise_Exported_NonEU_Annex1_thousandsUSD), log) %>%
  mutate(FDI_netinflow_perc_ofGDP = log(abs(FDI_netinflow_perc_ofGDP) + 0.00106616457) * sign(FDI_netinflow_perc_ofGDP)) %>%
  mutate(Electricity_from_renewables_perc_oftotal = log1p(Electricity_from_renewables_perc_oftotal)) %>%
  mutate(Merchandise_Exported_NonEU_Annex1_thousandsUSD = log1p(Merchandise_Exported_NonEU_Annex1_thousandsUSD))
  
  

 # combing log transformation values with the non-numeric data

imputed_data_non_numeric <- select(imputed_data, .imp, .id, Name, Year)
imputed_data_final_logged <-combined_df <- cbind(imputed_data_non_numeric, imputed_data_numeric_logged)
  
```



